Making Q-learning robust to time discretization

### 2019-01-18 paper/icml_drau/advup.tex
Commiting figures.

\subsection{Algorithms}
\label{subsec:algorithm}
To learn $V_{\theta}$ and $A_{\psi}$, variants of Q-learning for continuous and
discrete action spaces are used. In details, $V_{\theta}$ and $A_{\psi}$ are learnt
using approximate version of the following bellman equation, which holds for the true $A^\pi_\deltat$
and $V^\pi_\deltat$, with a greedy exploitation policy $\pi(s) = \text{argmax}_{a'}A^\pi_\deltat(s, a')$
\begin{align}
	V^{\pi}_\deltat(s) + \deltat A^{\pi}_\deltat(s, a) &= r\deltat + \gamma^{\deltat}  \E_{s'} V^{\pi}_\deltat(s')\label{eq:bellman_A}\\
	A^{\pi}_\deltat(s, \pi(s)) &= 0\label{eq:max_A}.
\end{align}
Maximization over actions is implemented exactly in the discrete actions case,
and approximated by a policy neural network $\pi_\phi(s)$, trained to maximize $A(s,
\pi_\phi(s))$ in the continuous case.  This is similar to what is done in
\cite{ddpg}.

Eq.~\eqref{eq:max_A}
is directly verified by $A_{\psi}$, owing to the reparametrization
$A_\psi(s, a) = \bar{A}_\psi(s, a) - \bar{A}_\psi(s, \pi(s))$, described
in~\ref{subsec:reparam}.  To approximately verify~\eqref{eq:bellman_A}, the
corresponding squared residual is minimized by an approximate gradient descent.
The update equations when learning from a transition $(s, a, r, s')$, either from
an exploratory trajectory or from a replay buffer~\cite{dqn} are
\begin{align}
	\delta Q_\deltat &\leftarrow A(s, a) \deltat - \left(r \deltat + \gamma^{\deltat} V(s') - V(s)\right)\\
	\theta_\deltat &\leftarrow \theta_\deltat + \eta^V_\deltat \partial_{\theta} V(s) \frac{\delta Q_\deltat}{\deltat}\\
	\psi_\deltat &\leftarrow \psi_\deltat + \eta^A_\deltat \partial_{\psi} A(s, a) \frac{\delta Q_\deltat}{\deltat}.
	\label{eq:approx_bellman_A}
\end{align} 
where $\eta$'s are learning rates.
Appropriate scalings for the learning rates $\eta^V_\deltat$ and $\eta^A_\deltat$ in term of $\deltat$ to obtain a well defined continuous limit are derived in the following section. \TODO{Change notations for the learning rates in the algorithm}
% As we will see in the following section,
% to obtain a well defined continuous time limit, the learning rates can be rewritten as
% $\eta^i_\deltat = \alpha^i \deltat$, where $\alpha^1$ and $\alpha^2$ are independant
% of $\deltat$.
\begin{algorithm}[ht]
	\caption{Deep Advantage Updating (Discrete actions)}
	\input{dauc.tex}
	\label{alg:dau}
\end{algorithm}

\subsection{Learning rates scalings}
\label{subsec:lr}
For the discrete algorithm to admit a continuous time limit, the trajectories
of parameters must converge to well defined trajectories as $\deltat$ goes to
$0$.  Intuitively, this means that optimization steps should not be too large,
in which case parameter trajectories could diverge in a single gradient step,
or too small, in which case the trajectories would converge to single points as
$\deltat$ goes to $0$.  This in turn imposes conditions on the scalings of the
learning rates of algorithms robust to changes of the tine discretization.

>>>>

\subsection{Algorithms}
\label{subsec:algorithm}
\begin{algorithm}[ht]
	\caption{Deep Advantage Updating (Discrete actions)}
	\input{dauc.tex}
	\label{alg:dau}
\end{algorithm}

To learn $V_{\theta}$ and $A_{\psi}$, variants of Q-learning for continuous and
discrete action spaces are used. In details, $V_{\theta}$ and $A_{\psi}$ are learnt
using approximate version of the following bellman equation, which holds for the true $A^\pi_\deltat$
and $V^\pi_\deltat$, with a greedy exploitation policy $\pi(s) = \text{argmax}_{a'}A^\pi_\deltat(s, a')$
\begin{align}
	V^{\pi}_\deltat(s) + \deltat A^{\pi}_\deltat(s, a) &= r\deltat + \gamma^{\deltat}  \E_{s'} V^{\pi}_\deltat(s')\label{eq:bellman_A}\\
	A^{\pi}_\deltat(s, \pi(s)) &= 0\label{eq:max_A}.
\end{align}
Maximization over actions is implemented exactly in the discrete actions case,
and approximated by a policy neural network $\pi_\phi(s)$, trained to maximize $A(s,
\pi_\phi(s))$ in the continuous case.  This is similar to what is done in
\cite{ddpg}.

Eq.~\eqref{eq:max_A}
is directly verified by $A_{\psi}$, owing to the reparametrization
$A_\psi(s, a) = \bar{A}_\psi(s, a) - \bar{A}_\psi(s, \pi(s))$, described
in~\ref{subsec:reparam}.  To approximately verify~\eqref{eq:bellman_A}, the
corresponding squared residual is minimized by an approximate gradient descent.
The update equations when learning from a transition $(s, a, r, s')$, either from
an exploratory trajectory or from a replay buffer~\cite{dqn} are
\begin{align}
	\delta Q_\deltat &\leftarrow A(s, a) \deltat - \left(r \deltat + \gamma^{\deltat} V(s') - V(s)\right)\\
	\theta_\deltat &\leftarrow \theta_\deltat + \eta^V_\deltat \partial_{\theta} V(s) \frac{\delta Q_\deltat}{\deltat}\\
	\psi_\deltat &\leftarrow \psi_\deltat + \eta^A_\deltat \partial_{\psi} A(s, a) \frac{\delta Q_\deltat}{\deltat}.
	\label{eq:approx_bellman_A}
\end{align} 
where $\eta$'s are learning rates.
Appropriate scalings for the learning rates $\eta^V_\deltat$ and $\eta^A_\deltat$ in term of $\deltat$ to obtain a well defined continuous limit are derived in the following section. \TODO{Change notations for the learning rates in the algorithm}
% As we will see in the following section,
% to obtain a well defined continuous time limit, the learning rates can be rewritten as
% $\eta^i_\deltat = \alpha^i \deltat$, where $\alpha^1$ and $\alpha^2$ are independant
% of $\deltat$.
\subsection{Learning rates scalings}
\label{subsec:lr}
\input{pendulum_fig.tex}
For the discrete algorithm to admit a continuous time limit, the trajectories
of parameters must converge to well defined trajectories as $\deltat$ goes to
$0$.  Intuitively, this means that optimization steps should not be too large,
in which case parameter trajectories could diverge in a single gradient step,
or too small, in which case the trajectories would converge to single points as
$\deltat$ goes to $0$.  This in turn imposes conditions on the scalings of the
learning rates of algorithms robust to changes of the tine discretization.

### 2019-01-18 paper/icml_drau/pendulum_fig.tex
Commiting figures.

%! TEX root = icml_drau.tex
\newcommand{\iw}{1cm}
\begin{figure*}[!ht]
	\begin{tabular}{c|c}
		DDPG & AU \\
		\hline
		\begin{tabular}{l|lllll}
			$\delta t = 0.01$ & \includegraphics[width=\iw]{figs_data/pendulum/0.01/ddpg_act_2.png} &\includegraphics[width=\iw]{figs_data/pendulum/0.01/ddpg_act_4.png} &\includegraphics[width=\iw]{figs_data/pendulum/0.01/ddpg_act_6.png} &\includegraphics[width=\iw]{figs_data/pendulum/0.01/ddpg_act_8.png} &\includegraphics[width=\iw]{figs_data/pendulum/0.01/ddpg_act_10.png} \\
			$\delta t = 0.005$ &\includegraphics[width=\iw]{figs_data/pendulum/0.005/ddpg_act_4.png} &\includegraphics[width=\iw]{figs_data/pendulum/0.005/ddpg_act_8.png} &\includegraphics[width=\iw]{figs_data/pendulum/0.005/ddpg_act_12.png} &\includegraphics[width=\iw]{figs_data/pendulum/0.005/ddpg_act_16.png} &\includegraphics[width=\iw]{figs_data/pendulum/0.005/ddpg_act_20.png} \\
			$\delta t = 0.002$ &\includegraphics[width=\iw]{figs_data/pendulum/0.002/ddpg_act_10.png} &\includegraphics[width=\iw]{figs_data/pendulum/0.002/ddpg_act_20.png} &\includegraphics[width=\iw]{figs_data/pendulum/0.002/ddpg_act_30.png} &\includegraphics[width=\iw]{figs_data/pendulum/0.002/ddpg_act_40.png} &\includegraphics[width=\iw]{figs_data/pendulum/0.002/ddpg_act_50.png} \\
			$\delta t = 0.001$ &\includegraphics[width=\iw]{figs_data/pendulum/0.001/ddpg_act_20.png} &\includegraphics[width=\iw]{figs_data/pendulum/0.001/ddpg_act_40.png} &\includegraphics[width=\iw]{figs_data/pendulum/0.001/ddpg_act_60.png} &\includegraphics[width=\iw]{figs_data/pendulum/0.001/ddpg_act_80.png} &\includegraphics[width=\iw]{figs_data/pendulum/0.001/ddpg_act_100.png} \\
			$\delta t = 0.0005$ &\includegraphics[width=\iw]{figs_data/pendulum/0.0005/ddpg_act_40.png} &\includegraphics[width=\iw]{figs_data/pendulum/0.0005/ddpg_act_80.png} &\includegraphics[width=\iw]{figs_data/pendulum/0.0005/ddpg_act_120.png} &\includegraphics[width=\iw]{figs_data/pendulum/0.0005/ddpg_act_160.png} &\includegraphics[width=\iw]{figs_data/pendulum/0.0005/ddpg_act_200.png}
		\end{tabular} &
		\begin{tabular}{lllll}
			 \includegraphics[width=\iw]{figs_data/pendulum/0.01/au_act_2.png} &\includegraphics[width=\iw]{figs_data/pendulum/0.01/au_act_4.png} &\includegraphics[width=\iw]{figs_data/pendulum/0.01/au_act_6.png} &\includegraphics[width=\iw]{figs_data/pendulum/0.01/au_act_8.png} &\includegraphics[width=\iw]{figs_data/pendulum/0.01/au_act_10.png} \\
			\includegraphics[width=\iw]{figs_data/pendulum/0.005/au_act_4.png} &\includegraphics[width=\iw]{figs_data/pendulum/0.005/au_act_8.png} &\includegraphics[width=\iw]{figs_data/pendulum/0.005/au_act_12.png} &\includegraphics[width=\iw]{figs_data/pendulum/0.005/au_act_16.png} &\includegraphics[width=\iw]{figs_data/pendulum/0.005/au_act_20.png} \\
			\includegraphics[width=\iw]{figs_data/pendulum/0.002/au_act_10.png} &\includegraphics[width=\iw]{figs_data/pendulum/0.002/au_act_20.png} &\includegraphics[width=\iw]{figs_data/pendulum/0.002/au_act_30.png} &\includegraphics[width=\iw]{figs_data/pendulum/0.002/au_act_40.png} &\includegraphics[width=\iw]{figs_data/pendulum/0.002/au_act_50.png} \\
			\includegraphics[width=\iw]{figs_data/pendulum/0.001/au_act_20.png} &\includegraphics[width=\iw]{figs_data/pendulum/0.001/au_act_40.png} &\includegraphics[width=\iw]{figs_data/pendulum/0.001/au_act_60.png} &\includegraphics[width=\iw]{figs_data/pendulum/0.001/au_act_80.png} &\includegraphics[width=\iw]{figs_data/pendulum/0.001/au_act_100.png} \\
			\includegraphics[width=\iw]{figs_data/pendulum/0.0005/au_act_40.png} &\includegraphics[width=\iw]{figs_data/pendulum/0.0005/au_act_80.png} &\includegraphics[width=\iw]{figs_data/pendulum/0.0005/au_act_120.png} &\includegraphics[width=\iw]{figs_data/pendulum/0.0005/au_act_160.png} &\includegraphics[width=\iw]{figs_data/pendulum/0.0005/au_act_200.png}
		\end{tabular}
	\end{tabular}
	\caption{Policies obtained by DDPG and AU at different macroscopic instants of training. Each image represents the policy learnt by the policy network, with $x$-axis representing angle, and $y$-axis angular velocity. The lighter the pixel, the closer to $1$ the action, the darker, the closer to $-1$.}
\end{figure*}

>>>>

%! TEX root = icml_drau.tex
\newcommand{\iw}{1cm}
\begin{figure*}[ht]
	\includegraphics[width=\textwidth]{figs_data/pendulum_fig.pdf}
	\caption{Policies obtained by DDPG and AU at different macroscopic instants of training. Each image represents the policy learnt by the policy network, with $x$-axis representing angle, and $y$-axis angular velocity. The lighter the pixel, the closer to $1$ the action, the darker, the closer to $-1$.}
\end{figure*}
